---
title: "Chapter 5: Dimensionality reduction techniques"
author: "Gonzalo de Quesada"
date: "4 de diciembre de 2017"
output: html_document
---

# Chapter 5: Dimensionality reduction techniques

The dataset was obtained from the United Nations Development Program.

```{r, dat}
setwd("C:/Users/D E L L/Documents/GitHub/IODS-project/Data")
human <- read.table("human.csv", header = TRUE, sep = ",")

str(human)
dim(human)
```

Now we vizualize the data using a correlation plot matrix

```{r,plot00}
library(GGally)
ggpairs(human)
library(corrplot)
humancor <- cor(human)
corrplot(humancor)
```

We can see a strong negative correlation between maternal mortality, econdary education ratio and years of expected education

# Principal component analysis (PCA)

```{r, pca}
pca_human <- prcomp(human)
s <- summary(pca_human)
s
pca_pr <- round(100*s$importance[2,], digits=1)
pca_pr
biplot(pca_human, choices = 1:2, cex = c(0.6, 1))
biplot(pca_human, choices = 1:2, cex = c(0.6, 0.9))
```

# Multiple correspondance analysis

We will use the dataset tea for this analysis. the data contains 300 observations of 36 variables with information about tea drinking habits

```{r, tea}
library(FactoMineR)
data("tea")
str(tea)
dim(tea)
```
```{r, mca}
library(dplyr)
library(tidyr)
library(ggplot2)
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")
tea_time <- select(tea, one_of(keep_columns))
gather(tea_time) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
```

We can see that most people use tea bags, have it alone, not at lunch, early grey and from a chain store. the consumption of sugar its split.

```{r, mca1}
mca <- MCA(tea_time, graph = FALSE)
summary(mca)
plot(mca, invisible = "ind", habillage = "quali")
plot(mca, invisible = "var", habillage = "quali")
```