## "Chapter 2. Regression and model validation"

Data

Our original data


```{r,read}
learning2014col <- read.csv("C:/Users/D E L L/Documents/GitHub/IODS-project/Data/learning2014.csv")
str(learning2014col)
```
```{r, dim}
dim(learning2014col)
```

the data have 7 variables:
Gender: It has 2 levels: M for male, F for female

Age: age of the students.

Attitude: attitude towards statistics.

Deep: deep approach learning and it is a combination of variables including: seeking meaning, relating ideas and use of evidence. This variables measure sto what extent the student has the intention to maximize understanding, with a true commitment to learning.

Stra: strategic approach and it is a combination of variables including: organized studying and time management. This variables measures to what extent the students organize their studying

Surf: surface approach learning and it is a combination of variables including: lack of purpose, unrelated memorising and syllabus-boundness.This variable measures to what extent the student memorizes without understanding, with a serious lack of personal engagement in the learning process.

Points: the exam points that students obtained. 

for a graphic overviwe we will use ggplot2 library because it creates scatter plots and correlation value. the mapping describes the aesthetic parameters and the colour between male and female.
```{r, library}
library(ggplot2)

library(GGally)
```

```{r, ggpairs}
p <- ggpairs(learning2014col, mapping = aes(col=gender), lower=list(combo=wrap("facethist", bins=20)))
p
```

using this model we choose three variables as explanatory variables (surf, stra, attitude) of the dependent variable (exam points)

```{r, lm}
model <- lm(Points ~ Attitude + stra + surf, learning2014col)
model
```
the output shows the t values, the standart error and p values.

the model with only one variable (attitude)

```{r, lmpoints}
model1 <- lm(Points ~ Attitude, learning2014col)
model1
```
## residual vs fitted values 
```{r, plot}
plot(model1, 1)
```

## normal qq-plot
```{r, plot1}
plot(model1, 2)
```

## residual vs lererage
```{r, plot2}
plot(model1, 5)
```

